---
title: "coAuthor-ENA"
author: "Zach"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Load packages
```{r message=F, warning=F, paged.print=T}

rm(list=ls()) #clear environment

library(rENA)
#library(ona)
#library(tma)
library(tidyverse) #for wrangling
library(lmerTest) #for hlms
library(ICC) #for testing clustering of observations
library(emmeans)  #for comparing subpopulations 
library(performance) #for regression diagnostics

```

# Prep data

## Read data
```{r}
data1 <- read.csv('ena_all_new_v2.csv',stringsAsFactors = FALSE) #read data
# 
#read metadata

meta_cr = read.csv("~/Rprojects/Yixin/CoAuthor - Metadata & Survey - Metadata (creative).csv",
                   stringsAsFactors = FALSE)

meta_arg = read.csv("~/Rprojects/Yixin/CoAuthor - Metadata & Survey - Metadata (argumentative).csv",
                   stringsAsFactors = FALSE)

meta_coauthor = bind_rows(meta_cr,meta_arg)


#load("~/Rprojects/Yixin/accum-300823.Rdata")
```


## Add metadata
```{r}
data1 = left_join(data1,meta_coauthor,by = c("worker_id","session_id"))
```



# Prep for ENA model
```{r }

units = data1[,c("session_id",
                 "worker_id")]


conversation = data1[,c("session_id",
                        "worker_id",
                        "sentSeq")]

codeCols = c(
  'compose',
  #'delete',
  'relocate',
  'reflect',
  'seekSugg',
  'acceptSugg',
  'dismissSugg',
  'lowModification',
  'highModification',
  'reviseSugg',
  'reviseUser',
  "cursorFwd",
  "cursorBwd",
  "cursorSelect",
  #"reopenSugg",
  "hoverSugg"
)

codes = data1[,codeCols]

#mask =

meta = data1[,c("genre",
                "highTemp",
                "ownershipMetadata",
                "prompt_code"
                )]

```

# Run ENA accumulation
```{r}
accum =
  ena.accumulate.data(
  units = units,
  conversation = conversation,
  codes = codes,
  metadata = meta,
  #mask = mask,
  window.size.back = "inf" # each line in the conversation can connect back to the first line--allows for more fine-grained view of connections w/i convos
)
```

# Run ENA dimensional reduction
```{r}
set = ena.make.set(accum)
```

# View space
```{r}
network = as.matrix(set$line.weights)
mean_network = colMeans(network)

network_mult = 0

p = ena.plot(set,title = "Overall Mean Network") %>%
  ena.plot.network(mean_network * network_mult,colors = "black")

p$plot
```

# Statistical tests

## Set up data and check data
```{r}
#names(set$points)
reg_data = set$points[,c(1:9)]
#glimpse(reg_data)
#table(reg_data$genre)
#t(table(reg_data$genre,reg_data$worker_id))
#summary(reg_data)
```

## Checking points
```{r}
ggplot(reg_data,aes(x = SVD1, y = SVD2)) + geom_point()
```

## Checking other groups

### genre
```{r}
ggplot(reg_data, aes(x = genre, y = SVD1)) + geom_boxplot()
ggplot(reg_data, aes(x = genre, y = SVD2)) + geom_boxplot()

```
### ownership
```{r}
ggplot(reg_data, aes(x = ownershipMetadata, y = SVD1)) + geom_boxplot()
ggplot(reg_data, aes(x = ownershipMetadata, y = SVD2)) + geom_boxplot()

```

### temperature
```{r}
ggplot(reg_data, aes(x = as.factor(highTemp), y = SVD1)) + geom_boxplot()
ggplot(reg_data, aes(x = as.factor(highTemp), y = SVD2)) + geom_boxplot()
```

## Clustering of observations
```{r}
ICCest(worker_id,SVD1,reg_data) #CI does not contain zero; significant
ICCest(worker_id,SVD2,reg_data) #CI does not contain zero; significant

#suggests multilevel models are appropriate for these data

ICCest(prompt_code,SVD1,reg_data) #CI does contains zero; not significant
ICCest(prompt_code,SVD2,reg_data) #CI contains zero; not significant
```

## Regression analysis

### SVD1
```{r}
#mod.x.1 = lmerTest::lmer(SVD1 ~ genre*ownershipMetadata*highTemp + (1|worker_id),data = reg_data)

#confint(mod.x.1)

#mod.x.2 = lmerTest::lmer(SVD1 ~ genre + ownershipMetadata + highTemp + (1|worker_id),data = reg_data)

#confint(mod.x.2)

#suggests only ownership is signficant

mod.x.3 = lmerTest::lmer(SVD1 ~ ownershipMetadata + (1|worker_id),data = reg_data)

summary(mod.x.3)

```
#### Check model
```{r}
check.x = check_model(mod.x.3, check = c("qq","normality","linearity", "homogeneity", "outliers","reqq")) #looks pretty good

check.x

check_outliers(mod.x.3,"mahalanobis")
```

#### Effect size
```{r}

demean = function(x){
  return(x - mean(x))
}

cohensd = function(diff_,x1,x2){
  num = diff_
  denom = sqrt((sum(demean(x1)^2) + sum(demean(x2)^2))/(length(x1) + length(x2) - 2))
  return(num/denom)
  
}

```


```{r}
diff_ = coefficients(mod.x.3)$worker_id$ownershipMetadatauser[1]
x1 = as.matrix(set$points$ownershipMetadata$user)[,"SVD1"]
x2 = as.matrix(set$points$ownershipMetadata$api)[,"SVD1"]

cohensd(diff_ = diff_,x1,x2)


```

### SVD2

```{r}
# mod.y.1 = lmerTest::lmer(SVD2 ~ genre*ownershipMetadata*highTemp + (1|worker_id),data = reg_data)
# 
# summary(mod.y.1)
# 
# 
# mod.y.2 = lmerTest::lmer(SVD2 ~ genre + ownershipMetadata + highTemp  + (1|worker_id),data = reg_data)
# 
# summary(mod.y.2)

mod.y.3 = lmerTest::lmer(SVD2 ~ genre  + (1|worker_id),data = reg_data)

summary(mod.y.3)

```
#### check model
```{r}
check.y = check_model(mod.y.3, check = c("qq","normality","linearity", "homogeneity", "outliers","reqq")) #looks pretty good

check.y

check_outliers(mod.y.3,"mahalanobis")
```

#### Effect size (genre)
```{r}
diff_ = coefficients(mod.y.3)$worker_id$genre[1]
x1 = as.matrix(set$points$genre$creative)[,"SVD2"]
x2 = as.matrix(set$points$genre$argumentative)[,"SVD2"]

cohensd(diff_,x1,x2)
```


# ENA plots

## Mean networks (ownership)
```{r}
user_pts = as.matrix(set$points$ownershipMetadata$user)
api_pts = as.matrix(set$points$ownershipMetadata$api)

user_net = colMeans(as.matrix(set$line.weights$ownershipMetadata$user))
api_net = colMeans(as.matrix(set$line.weights$ownershipMetadata$api))

plot_user = ena.plot(set, scale.to = "network", title = "User")  %>%
       #ena.plot.points(points = creat_pts, colors = c("blue"))  %>%
       ena.plot.group(point = user_pts,
                      colors =c("blue"), confidence.interval = "none")  %>%
       ena.plot.network(network = user_net,colors = c("blue") )


plot_api = ena.plot(set, scale.to = "network", title = "Api")  %>%
      # ena.plot.points(points = arg_pts, colors = c("red"))  %>%
       ena.plot.group(point = api_pts,
                      colors =c("red"), confidence.interval = "none")  %>%
       ena.plot.network(network = api_net,colors = c("red") )

plot_user$plot
plot_api$plot

```
## Network subtraction (ownership)
```{r }

net_mult = 3

plot_sub = ena.plot(set, scale.to = "network", title = "API vs User")  %>%
       ena.plot.group(point = user_pts,
                      colors =c("blue"), confidence.interval = "box")  %>%
       ena.plot.group(point = api_pts,
                      colors =c("red"), confidence.interval = "box") %>% 
       ena.plot.network(network = (user_net - api_net) * net_mult,colors = c("blue","red") )
plot_sub$plot
```
## Mean networks (genre)
```{r}
creat_pts = as.matrix(set$points$genre$creative)
arg_pts = as.matrix(set$points$genre$argumentative)

creat_net = colMeans(as.matrix(set$line.weights$genre$creative))
arg_net = colMeans(as.matrix(set$line.weights$genre$argumentative))

plot_cr = ena.plot(set, scale.to = "network", title = "Creative")  %>%
       #ena.plot.points(points = creat_pts, colors = c("blue"))  %>%
       ena.plot.group(point = creat_pts,
                      colors =c("blue"), confidence.interval = "none")  %>%
       ena.plot.network(network = creat_net,colors = c("blue") )


plot_arg = ena.plot(set, scale.to = "network", title = "Argumentative")  %>%
      # ena.plot.points(points = arg_pts, colors = c("red"))  %>%
       ena.plot.group(point = arg_pts,
                      colors =c("red"), confidence.interval = "none")  %>%
       ena.plot.network(network = arg_net,colors = c("red") )

plot_cr$plot
plot_arg$plot

```

## Network subtraction (genre)
```{r }

net_mult = 5

plot_sub = ena.plot(set, scale.to = "network", title = "Creative vs Argumentative")  %>%
       ena.plot.group(point = creat_pts,
                      colors =c("blue"), confidence.interval = "box")  %>%
       ena.plot.group(point = arg_pts,
                      colors =c("red"), confidence.interval = "box") %>% 
       ena.plot.network(network = (creat_net - arg_net) * net_mult,colors = c("blue","red") )
plot_sub$plot
```